---
title: "Counting Characters"
description: |
  Using python and R to improve my typing experience.
author: Jannik Buhr
draft: false
date: 11-08-2020
output:
  distill::distill_article:
    toc: true
    self_contained: false
---

```{r setup, include=FALSE}
library(reticulate)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE)
```

## The Why

In this post I am doing something slightly unusual.
I am using python (in an Rmarkdown document of course)
to count the key utilization of R and Rmarkdown files.
The goal is twofold:

Firstly, I want to get better at python, beyond
theoretical studies and watching videos, so I have
to practice.
Secondly, I recently acquired a new keyboard
(the [moonlander](https://www.zsa.io/moonlander/)) and
while I tried to stick to a layout that makes
transitioning between my laptop and the keyboard
easy, I still have some space left on on of my symbols layer
that I want to populate with the most used keys.
And of course we will be doing this based on data!
So, without further ado, let's jump in.

## The How

I wasn't sure how many R files I had in my projects folder,
so I resorted to python's generator expressions to make sure
to not load everything into memory at once.

```{python}
from glob import iglob
from collections import Counter
from functools import reduce

path = "/home/jannik/Documents/projects/rstats"

def get_count(p):
  try:
    with open(p, 'r') as f:
      cs = [c for x in f.readlines()
              for c in x]
      return {char : cs.count(char) for char in set(cs)}
  except:
    return {'a':0}

def add_counts(d1, d2):
  result = Counter(d1) + Counter(d2)
  return result

```

Using functional programming techniques, this process
becomes rather compact.
A map of character counts per file is reduced to a single
dictionary of counts.

```{python}
files = iglob(path + "/*/**/*.R", recursive = True)
counts = map(get_count, files)
result = dict(reduce(add_counts, counts))
```


## The Result

And now back to R for visualization.
The `reticulate` package translates python dictionaries
to named lists in R, so we are good to go.

```{r}
library(tidyverse)

res <- enframe(py$result) %>% 
  mutate(value = unlist(value)) %>% 
  arrange(desc(value))

res %>% 
  mutate(name = case_when(
    name == " " ~ "space",
    name == "\n" ~ "enter",
    TRUE ~ name
  )) %>%
  filter(!(tolower(name) %in% c(letters, "space", "enter", 0:9))) %>% 
  slice_max(value, n = 20) %>% 
  mutate(name = fct_reorder(name, value)) %>% 
  ggplot(aes(value, name, fill = value)) +
  geom_col(color = "black", width = 0.8) +
  geom_text(aes(label = name),
            hjust    = -1,
            fontface = "bold") +
  scale_fill_viridis_c() +
  scale_x_continuous(expand = expand_scale(c(0.01, 0.05))) +
  scale_y_discrete(labels = NULL) +
  guides(fill = "none") +
  theme_minimal() +
  theme(axis.ticks.y = element_blank()) +
  labs(y       = "",
       x       = "count",
       caption = "excluding letters and numbers")
```

The overwhelming result: I need a better place
for the minus sign! Not because I am subtracting
numbers all the time, but because it is also
in the assignment operator `<-`, which can 
be inserted in RStudio using Alt+Minus.

[This](https://configure.ergodox-ez.com/moonlander/layouts/bdZ6A/latest/0) is
what my layout currently looks like.
